Modelling potential environmental impacts of science activity in Antarctica

### 2025-12-17 paper/paper.qmd
removed suggestion of figure of largest site
latest version of paper

## Introduction
**FRASER MATERIAL STARTS**

Antarctica is often considered a vast, desolate continent devoid of human activity.
However, there are small, scientifically significant areas where substantial, regular human activity occurs.
Globally unique ice-free areas within Antarctica, such as the McMurdo Dry Valleys, are of particular scientific interest and so a significant amount of activity occurs in this specially managed area.
Human activity in Antarctica is highly dependent on the infrastructure that supports people to survive in this environment. 
Outside of a reduced period of activity caused by COVID-19, human activity on the continent continues to grow [@Tejedo2024] and so the actual, and future potential impact of these activities continues to grow. 
The McMurdo Dry Valleys are close to the largest research base on the continent, McMurdo Station, and the pressures on this unique environment from human activity are increasing. 
Understanding the scope and implications of human activity in key parts of Antarctica is critical to supporting environmental management of the continent.

@@

More recent research has focused either on a continental-scale picture of human activity [@Leihy2020], or detailed mapping of the infrastructural footprint and impacts that the numerous Antarctic stations have on the continent [@Brooks2018,@Brooks2019].
These approaches tackle the challenge of understanding human activity from contrasting spatial perspectives. 
The present research aims to complement these two techniques, by using high resolution mapping of human activity in the dry valleys to extrapolate how scientists might navigate around these landscapes.

@@

## Discussion {#sec-discussion}
To cover:

+ Figure of the largest study site?
+ Limitations of the method:
  + Data concerns, esp. opportunistic nature of the GPS data
  + Would higher resolution network be useful?
  + Possible limitations of a regularly spaced lattice: extend to a lattice based on e.g., TIN?
  + Reliance on shortest paths as a guide to behaviour
  + Some minor issues regarding the difference between directed and undirected graphs
+ Relevance of radius limits on betweenness: the implicit model of movement in this approach (short horizons, no existing paths, 'keep going for half an hour in this general direction...')
+ Question of embedding paths, i.e., locking in more damage to some parts of the landscape vs others

>>>>

## Introduction
Antarctica is often considered a vast, desolate continent devoid of human activity.
However, there are small, scientifically significant areas where substantial, regular human activity occurs.
Globally unique ice-free areas within Antarctica, such as the McMurdo Dry Valleys, are of particular scientific interest and so a significant amount of activity occurs in this specially managed area.
Human activity in Antarctica is highly dependent on the infrastructure that supports people to survive in this environment. 
Outside of a reduced period of activity caused by COVID-19, human activity on the continent continues to grow [@Tejedo2024] and so the actual, and future potential impact of these activities continues to grow. 
The McMurdo Dry Valleys are close to the largest research base on the continent, McMurdo Station, and the pressures on this unique environment from human activity are increasing. 
Understanding the scope and implications of human activity in key parts of Antarctica is critical to supporting environmental management of the continent.

@@

More recent research has focused either on a continental-scale picture of human activity [@Leihy2020], or detailed mapping of the infrastructural footprint and impacts that the numerous Antarctic stations have on the continent [@Brooks2018; @Brooks2019].
These approaches tackle the challenge of understanding human activity from contrasting spatial perspectives. 
The present research aims to complement these two techniques, by using high resolution mapping of human activity in the dry valleys to extrapolate how scientists navigate in these landscapes, or could potentially navigate in these landscapes.

@@

## Discussion {#sec-discussion}
+ Limitations of the method:
  + Data concerns, esp. given improvements in  GPS data collection since 2016/18
  + Would higher resolution network be useful? (we think probably not)
  + Possible limitations of a regularly spaced lattice: extend to a lattice based on e.g., TIN?
  + Reliance on shortest paths as a guide to behaviour
  + Some minor issues regarding the difference between directed and undirected graphs (to check using `networkx`)
+ Relevance of radius limits on betweenness: the implicit model of movement in this approach (short horizons, no existing paths, 'keep going for half an hour in this general direction...')
+ Question of embedding paths, i.e., locking in more damage to some parts of the landscape vs others (this leads to the question in conclusion re method for less damaging plans)
